---
title: "Covid19 vaccine study Fall2021"
author: "Martha Leal Olmedo"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 5
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
TheData <-read.csv(file="FallFinalData.csv",header=TRUE) %>% select(-"X.1", -"X")
Covid91 <- TheData %>% mutate(across(where(is.character),as.factor)) 
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males
```{r}
Males <- filter(Covid91, sex == "Man")
```

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")

```

From the two bar charts above, We can see that the males who took the placebo got more sick than the males who took the Covid19 drug. This indicates that the drug is effective for the males group.

### Numerical Summaries  

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

The information above shows that 33.11% of the sick males had the treatment, and that 66.89% of those males who got sick took the placebo. This is a two to one difference and this lets us know that the treatment is working.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Malesmod <- fisher.test(table1)
```

According to the Chi squared test above, it shows us that the p-value is very small, it is (2.2e-16). This indicates that we need to reject the null hypothesis and that there is a difference between the males the took the drug and the males that took the placebo.


### Calculating Effectiveness 
```{r}
MaleEstimate <- Malesmod$estimate
Malelow <- Malesmod$conf.int[1]
Malehigh <- Malesmod$conf.int[2]

Malelowprob <-Malelow / (1 + Malelow)
Maleloweffect <- Malelowprob * 100

Malehighprob <- Malehigh / (1+ Malehigh)
Malehigheffect <- Malehighprob * 100

Maleprob <- MaleEstimate / (1+ MaleEstimate)
MaleEffect <- Maleprob * 100
options(digits = 1)
```

The effectiveness of the covid91 vaccine for Males is `r MaleEffect`. The confidence interval for males is `r Maleloweffect` to `r Malehigheffect `


# Females
```{r}
Females <- filter(Covid91, sex == "Woman")
```

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

From the two bar chart above, we can see that females that took the covid19 drug got more sick than the females that took the placebo. This indicates that the drug is not effective for the Females group.

### Numerical Summaries  

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

The information above shows that the females that got sick, 58% of them had taken the vaccine, and that 42% of the females sick took the placebo. Based on the information that was given, we can see that the vaccine seems to be ineffective for the females, so the vaccine is not working as good as it should.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Femalesmod <- fisher.test(table1)
```

The Chi squared test above shows us that the p value is 6e-04. This indicates that the p value is small, so we reject the hypothesis. In this case it shows that the data is different in a negative way, as in the drug is not effective compared to the placebo.


 

### Calculating Effectiveness  
```{r}
FemaleEstimate <- Femalesmod$estimate
Femalelow <- Femalesmod$conf.int[1]
Femalehigh <- Femalesmod$conf.int[2]

Femalelowprob <-Femalelow / (1 + Femalelow)
Femaleloweffect <- Femalelowprob * 100

Femalehighprob <- Femalehigh / (1+ Femalehigh)
Femalehigheffect <- Femalehighprob * 100

Femaleprob <- FemaleEstimate / (1+ FemaleEstimate)
FemaleEffect <- Femaleprob * 100
options(digits = 1)
```

The effectiveness of the covid91 vaccine for Females is `r FemaleEffect`. The confidence interval for Females is `r Femaleloweffect` to `r Femalehigheffect `


# Drug Users

# LGBTQ

# Conclusions and recommendations

