---
title: "Covid19 vaccine study Fall2021"
author: "Martha Leal Olmedo"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 5
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
TheData <-read.csv(file="FallFinalData.csv",header=TRUE) %>% select(-"X.1", -"X")
Covid91 <- TheData %>% mutate(across(where(is.character),as.factor)) 
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males
```{r}
Males <- filter(Covid91, sex == "Man")
```

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")

```

From the two bar charts above, We can see that the males who took the placebo got more sick than the males who took the Covid91 drug. This indicates that the drug is effective for the males group.

### Numerical Summaries  

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

The information above shows that 33.11% of the sick males had the treatment, and that 66.89% of those males who got sick took the placebo. This is a two to one difference and this lets us know that the treatment is working.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Malesmod <- fisher.test(table1)
```

According to the Chi squared test above, it shows us that the p-value is very small, it is (2.2e-16). This indicates that we need to reject the null hypothesis and that there is a difference between the males the took the drug and the males that took the placebo.


### Calculating Effectiveness 
```{r}
MaleEstimate <- Malesmod$estimate
Malelow <- Malesmod$conf.int[1]
Malehigh <- Malesmod$conf.int[2]

Malelowprob <-Malelow / (1 + Malelow)
Maleloweffect <- Malelowprob * 100

Malehighprob <- Malehigh / (1+ Malehigh)
Malehigheffect <- Malehighprob * 100

Maleprob <- MaleEstimate / (1+ MaleEstimate)
MaleEffect <- Maleprob * 100
options(digits = 1)
```

The effectiveness of the covid91 vaccine for Males is `r MaleEffect`. The confidence interval for males is `r Maleloweffect` to `r Malehigheffect `


# Females
```{r}
Females <- filter(Covid91, sex == "Woman")
```

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

From the two bar chart above, we can see that females that took the covid19 drug got more sick than the females that took the placebo. This indicates that the drug is not effective for the Females group.

### Numerical Summaries  

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

The information above shows that the females that got sick, 58% of them had taken the vaccine, and that 42% of the females sick took the placebo. Based on the information that was given, we can see that the vaccine seems to be ineffective for the females, so the vaccine is not working as good as it should.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Femalesmod <- fisher.test(table1)
```

The Chi squared test above shows us that the p value is 6e-04. This indicates that the p value is small, so we reject the hypothesis. In this case it shows that the data is different in a negative way, as in the drug is not effective compared to the placebo.


 

### Calculating Effectiveness  
```{r}
FemaleEstimate <- Femalesmod$estimate
Femalelow <- Femalesmod$conf.int[1]
Femalehigh <- Femalesmod$conf.int[2]

Femalelowprob <-Femalelow / (1 + Femalelow)
Femaleloweffect <- Femalelowprob * 100

Femalehighprob <- Femalehigh / (1+ Femalehigh)
Femalehigheffect <- Femalehighprob * 100

Femaleprob <- FemaleEstimate / (1+ FemaleEstimate)
FemaleEffect <- Femaleprob * 100
options(digits = 1)
```

The effectiveness of the covid91 vaccine for Females is `r FemaleEffect`. The confidence interval for Females is `r Femaleloweffect` to `r Femalehigheffect `


# Drug Users
```{r}
DrugUsers <- filter(Covid91, DrugUser == "yes")
```

```{r}
barchartGC(~infected + treatment,data=DrugUsers)
barchartGC(~infected + treatment,data=DrugUsers, type="percent")

```

From the two bar chart above, we can see that Drug users that took the placebo got more sick than the drug users that took the covid91 drug, but we can also see that drug users that took the placebo appear to be healthier than the drug users that took the covid91 drug, This indicates that the vaccine is preventing drug users from getting sick but shows that they are less healthier than the ones who took the placebo. So in conclusion the vaccine is working but it needs some improvement.


### Numerical Summaries  

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUsers)
rowPerc(table1)
colPerc(table1)
```

The information above shows that, for sick drug users that took the drug is 33%, and the sick drug users that took the placebo was 67%. so based on the information it shows us that the vaccine seems to be working so the vaccine is ineffective.



## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
DrugUsersmod <- fisher.test(table1)
```

The Chi squared test above shows us that the p value is 6e-06. This indicates that the p value is small, so we reject the hypothesis. So in conclusion the vaccine for Drug users is ineffective, it does not work. 


### Calculating Effectiveness 
```{r}
DrugUsersEstimate <- DrugUsersmod$estimate
Druglow <- DrugUsersmod$conf.int[1]
Drughigh <- DrugUsersmod$conf.int[2]

Druglowprob <-Druglow / (1 + Druglow)
Drugloweffect <- Druglowprob * 100

Drughighprob <- Drughigh / (1+ Drughigh)
Drughigheffect <- Drughighprob * 100

Drugprob <- DrugUsersEstimate / (1+ DrugUsersEstimate)
DrugUserEffect <- Drugprob * 100
options(digits = 1)
```

The effectiveness of the covid91 vaccine for DrugUser is `r DrugUserEffect`. The confidence interval for DrugUser is `r Drugloweffect` to `r Drughigheffect `


# LGBTQ
```{r}
LGBTQ <- filter(Covid91, LGBTQ == "gay")
```

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

According to the bar charts above, it shows that the LGBTQ does not benefit from taking this drug. In the sick group, had taken the drug.

### Numerical Summaries  
```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

According to the information above, it shows us that the LGBTQ should not consider taking this drug, as 71% of them who were sick had taken the drug and 29% sick people took the placebo.



## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
LGBTQmod <- fisher.test(table1)
```

The Chi squared test shows us that the p value is 6e-06, which is very small. This indicates that we need to reject the null hypothesis and that there is a difference between the LGBTQ who took the placebo and those who took the drug. In this case the drug is ineffective, so there were more people who got infected while on the drug.


### Calculating Effectiveness 
```{r}
LGBTQEstimate <- LGBTQmod$estimate
LGBTQlow <- LGBTQmod$conf.int[1]
LGBTQhigh <- LGBTQmod$conf.int[2]

LGBTQlowprob <-LGBTQlow / (1 + LGBTQlow)
LGBTQloweffect <- LGBTQlowprob * 100

LGBTQhighprob <- LGBTQhigh / (1+ LGBTQhigh)
LGBTQhigheffect <- LGBTQhighprob * 100

LGBTQprob <- LGBTQEstimate / (1+ LGBTQEstimate)
LGBTQEffect <- LGBTQprob * 100
options(digits = 1)
```

The effectiveness of the covid91 vaccine for LGBTQ community is `r LGBTQEffect`. The confidence interval for LGBTQ commnuity is `r LGBTQloweffect` to `r LGBTQhigheffect `



# Conclusions and recommendations
In conclusion the effectiveness of the covid91 for males was 67.2, for the females it was 40.4, for the DrugUsers it was 64.7, and for the LGBTQ community it was 28.6. From this information we can see that the LGBTQ community had the least ineffective vaccine, we can see that the most effective vaccine was the males vaccine. I would recommend the LGBTQ community to not take the vaccine due to it being ineffective.
